---
title: "Big Data Science - Assignment 1"
author: "Arnau Casanova & Alex Flores"
date: '`r Sys.Date()`'
output:
  html_notebook: default
  pdf_document: default
subtitle: EEG Analysis
---

# Problem Definition
## Case: Does the brain works in the same way in front of a visual stimulus with alcoholic and non-alcoholic persons?
  
  With this experiment we will see if there is a significative diference between alcoholic and non-alcoholic users that are connected to 64 sensors within three type of visual experiments.
 

```{r setoption, echo=FALSE,warning=FALSE,message=FALSE}
install.packages("stringr")
library(ggplot2)
```

> The main **goal** is to check wheter a user is alcoholic or non-alcoholic through their signal pattern


#Step 3 - Data Acquisition
## Data Understanding

##How to load the data?
  
To load our data we will use the manual download method of the full egg tar, we just need 
to unzip the full tar and then unzip the selected users to deal with, also we need to
set the directories of the respective users readings, the process file function will
do all the job to manage the data in the way we want.

```{r,warning=FALSE,message=FALSE,results='hide'}
#Process file function

getwd()

user <- ""
state <- ""
trial <- ""
expType <- ""
sensors <- list()
userValues <- list()


#Step 4: 

process_file <- function(filename){
  conn <<- file(filename,open="r")
  linn <-readLines(conn)
  
  userAux <- str_sub(linn[1],3,13)
  stateAux <- str_sub(linn[1],6,6)
  expTypeA <- str_split_fixed(linn[4], ",", n=2)
  trialA <- expTypeA[2]
  expTypeA <- expTypeA[1]
  expTypeA <- str_split_fixed(expTypeA, "#", n=2)
  expTypeA <- expTypeA[2]
  expTypeA <- str_trim(expTypeA, side = c("left"))
  
  user <<- userAux
  print(user)
  
  state <<- stateAux
  print(state)
  
  expType <<- expTypeA
  print(expType)
  
  trial <<- trialA
  print(trial)

  sensorsIndex <- 1;
  valuesIndex <- 1;
  
  
  for (i in 5:length(linn)){
    firstChar <- str_split_fixed(linn[i], " ", n=4)
    firstChar <- firstChar[1]
    if(firstChar == "#"){
      testLinia <- linn[i]
      testLinia <- str_split_fixed(testLinia, " ", n=4)
      testLinia <- testLinia[2]
      sensors[sensorsIndex] <<- testLinia
      sensorsIndex <- sensorsIndex + 1
    }
    else{
      linia <- linn[i]
      value <- str_split_fixed(linia, " ", n=4) 
      value <- value[4]
      userValues[valuesIndex] <<- value
      valuesIndex <- valuesIndex + 1
    }
  }
}
```

##How to save the data?

After unziping the full data set and selecting specific users and processing those files,
three txt files will be created, always specifing the path, this is because the structure
of the taken data is not the way we want, the first one is a file with all the data of the
experiment and the others are the alcholic and non-alcoholic data sets, this file process may
take many minutes since the quantity of data we are managing is really big.

##In which format the data needs to be keept?

As every file has 16.000 rows we need to think a method to keep the data in a clear format,
observing that the sample is going always from 1 to 256, in order to avoid that noisy repetition we will structure each file with 256 rows, and the columns of the result will be: the user id, the state of the user(alcoholic/non-alcoholic), the experiment type, the sample, the trial and all the 64 sensors.


##How many data we want to deal with?

The objective of the assignment is to analyze differences between alcoholic and non-alcoholic,
we could take the full data set but it would be a lot of data, in our case we will analyze
two random users (alcoholic and non-alcholic) and we will compare the readings of them 
in order to extract some conclusions.

The following code will do the construction of the three files, using each iteration the function defined below:
```{r,warning=FALSE,message=FALSE,results='hide'}
sensorsRow <- list()
userInfo <- list()
userInfo[1] <- "USER"
userInfo[2] <- "STATE"
userInfo[3] <- "EXP_TYPE"
userInfo[4] <- "SAMPLE"
userInfo[5] <- "TRIAL"


for(q in 1: 2){
  ##fileList = EEG files path
  ##connFile = connection to specific output file(alcoholic/non-alcoholic)
  ##connFileAll = connection to general output file
  
  if (q==1) {
    setwd("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/assignmentArchives/co2a00004447")
    fileList <- list.files("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/assignmentArchives/co2a00004447")
    connFile <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/alcoholicTable.txt", 'a')  
    connFileAll <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/bigTable.txt", 'a')
    } 
  else{ 
    setwd("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/assignmentArchives/co2c0000337")
    fileList <- list.files("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/assignmentArchives/co2c0000337")
    connFile <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/userTable.txt", 'a')  
    connFileAll <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/bigTable.txt", 'a')
    }
    
    
    for (x in 1: length(fileList)){
  
      fileList[x]
      process_file(fileList[x])
  
      if(state == "a"){
        state <- "1"
      }else{
        state <- "0"
       }
  
      if (x == 1){
        for (i in 1: length(sensors)){
          sensors[i] <- toString(sensors[i])
        }
        header <- c(userInfo, sensors)
        header <- paste(unlist(header), collapse= " ")
        write(header, connFile, sep = " ")
        write(header, connFileAll, sep = " ")
      }
      for(i in 1: 256){
        indexAux <- 0
        for(j in 6: 69){
          sensorsRow[1] <- user
          sensorsRow[2] <- state
          expType <- str_replace_all(expType, " ", "")
          sensorsRow[3] <- expType
          sensorsRow[4] <- toString(i)
          trial <- str_replace_all(trial, " ", "")
          sensorsRow[5] <- trial
          sensorsRow[j] <- toString(userValues[i+indexAux])
          indexAux <- indexAux + 256
        }
      insertRow <- paste(unlist(sensorsRow), collapse=" ")
      write(insertRow, connFile, sep = " ")
      write(insertRow, connFileAll, sep = " ")
      }
    }
  close(connFileAll)
  close(connFile)
}
fileGeneral <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/bigTable.txt")
fileA <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/alcoholicTable.txt")
fileC <- file("C:/Users/Usuari/Documents/UNI/4/4_2/BigDataScience/assignment1/userTable.txt")
dfAlc <- read.table(fileA, header = TRUE, sep = " ")
dfCon <- read.table(fileC, header = TRUE, sep = " ")
df <- read.table(fileGeneral, header = TRUE, sep = " ")

```

After cleaning the dataset and keeping it in a tidy format, we can show the summary of the entire set of trials for each user:

```{r}
##120 trial control user summary
summary(dfCon)


##120 trial alcoholic user summary
summary(dfAlc)

#To keep the data more able to analyse we focus in one channel
#Control
summary(dfCon$Y)

#Alcoholic
summary(dfAlc$Y)

```
At this first and fast look we can clearly see the diference in the intervals min-max, where alcoholic users have significative lower and higher values, so there is a kind of inestability on alcoholic user compared to control users.

Also we can have a look to the mean and median values
```{r}
#Control
meanCon <- mean(dfCon$FP1)
medCon <- median(dfCon$FP1)
meanCon
medCon
#Alcoholic
meanAlc <- mean(dfAlc$F8)
medAlc <- median(dfAlc$FP1)
meanAlc
medAlc
```

#Step 4 - Data Exploration
## Do you detect any problem with your data?

Yes, in some files there is a fourth paradigm that we had not specified or taken care of before, that inconsistency lead us to discard the information from that file, in order to mantain our data in a clean format we remove every file that contains data that is out of our specificaction. If we take a look at the mean and the median we can also see a significative difference knowing the quantity of data we are dealing with

Also, in the process of the file, we had to take care on a whitespace between the trial and a whitespace between the paradigm because the read.table function identified columns by whitespace, so we had to strim every little detail that affected the process to obtain a confortable data to work with.

##Do the users have the same number of samples?
  
Not all the users have the full set of samples, but this is not a problem because the user still have a large set of files, so we can analyze the data even if one or two samples are missing, in a set with that bunch of data having one less sample is not very signicative, so we have many data to analyze anyway.


##4.1. Exercise 1 - Represent the 'FP1' channel (first one).

```{r,warning=FALSE,message=FALSE}
#Control
ggplot(dfCon, aes(x=SAMPLE, y=FP1)) + geom_point()  + facet_grid(. ~ EXP_TYPE)

#Alcoholic
ggplot(dfAlc, aes(x=SAMPLE, y=FP1)) + geom_point() + facet_grid(. ~ EXP_TYPE)
```

At this points, as the min-max, mean and median showed, we can see a significative abundance of the control user with the value of the sensor near zero, in the alcoholic user case, most of cases the value is far from zero.

##4.2. Represent the 'FP1' channel as well as the next three
  


```{r,warning=FALSE,message=FALSE}
#Control
ggplot(dfCon, aes(x=SAMPLE, y=FP2))+geom_point()+ facet_grid(. ~ EXP_TYPE)

#Alcoholic
ggplot(dfAlc, aes(x=SAMPLE, y=FP2))+geom_point()+ facet_grid(. ~ EXP_TYPE)

mean(dfAlc$FP2)
mean(dfCon$FP2)

#Control
ggplot(dfCon, aes(x=SAMPLE, y=F7))+geom_point() + facet_grid(. ~ EXP_TYPE)

#Alcoholic
ggplot(dfAlc, aes(x=SAMPLE, y=F7))+geom_point()+ facet_grid(. ~ EXP_TYPE)

mean(dfAlc$F7)
mean(dfCon$F7)

#Control
ggplot(dfCon, aes(x=SAMPLE, y=F8))+geom_point() + facet_grid(. ~ EXP_TYPE)

#Alcoholic
ggplot(dfAlc, aes(x=SAMPLE, y=F8))+geom_point() + facet_grid(. ~ EXP_TYPE)

mean(dfAlc$F8)
mean(dfCon$F8)

```
After observing the different channels we can see more or less the same, the mean of the alcoholic users is always greater, acording to the greater maximum and minimum values we saw before


##4.4. Are there any outliers? First you need to define what is an outlier?
An outlier is data that is bigger or smaller than the nearest points, for example if in sample 1 the value is 1 and in the sample 2 the value is 100 or -100, we can say there is an outlier, since the next value is very far, in our case we can't see any outliers.



# Step 5 - Data Cleaning

##Is our data in tidy format?

The data we have taken wasn't in a tidy format, after the transformation that we applied we can assume that yes, it is in a tidy format, we set the variables in columns, we keep the user observations in rows and we have different metadata in different table, one for alcoholic and one for non-alcoholic, we also have a general table with all the data.


Step 5
##Exercise: Clean the data set
```{r,warning=FALSE,message=FALSE}
for(a in 6: 69){
  meanCol <- mean(col)
  dfTest <- dfAlc[a][1]
  for(z in 1: length(col)){
    col[z] <- col[z]-meanCol
  }
  dfAlc[a][1] <- col
}
```
#Exercise: Obtain the quantiles
