---
title: "Big Data Science - Assignment 2"
author: "Arnau Casanova & Alex Flores"
date: "`r Sys.Date()`"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
subtitle: EEG Analysis
---


PCA
Load the dataset

```{r}
library(data.table)
library(magrittr)

mydata<-read.table("C:/Users/Arnau/Documents/BigDataScience/assignment1/bigTable.txt", header=TRUE )

```

Check available variables (predictors)
```{r}
colnames(mydata)
```



Take a look at the structure

```{r}
mydata$STATE <-as.factor(mydata$STATE)
str(mydata)
```
Conversion of categorical variables
```{r}

#Convert categorical variables
library(dummies)
new_my_data <- dummy.data.frame(mydata, names = c("USER", "STATE", "EXP_TYPE"))
str(new_my_data)
```
PCA
```{r}
prin_comp <- prcomp(new_my_data, scale. = T)
names(prin_comp)
```
Mean and sd of variables
```{r}
prin_comp$center
prin_comp$scale
prin_comp$rotation
```
Principal component score vectors in 1729x91 dimension
```{r}
dim(prin_comp$x)
```
Plot the resultant principal compoents
```{r}
biplot(prin_comp, scale = 0)
```
Standard deviation of each principal compoent
```{r}
std_dev <- prin_comp$sdev
```
Variance
```{r}
pr_var <- std_dev^2
```
Check variance of first 10 components, higher value higher will be the information of these components
```{r}
pr_var[1:10]
```
Compute the proportion of variance explained by each component. Divide variance by sum total variance:
```{r}
prop_varex <- pr_var/sum(pr_var)
prop_varex[1:20]

```
1st component 31.9%, 2nd 15.33% and so on. How many components for the modeling? use scree plot to se variability in the data.
```{r}
#scree plot
plot(prop_varex, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             type = "b")

```
So 30 components explains around 80sum% variance in the data set. We had 67 predictors at the start, now 30.

```{r}
sum(pr_var[1:30])
```

Check the cumulative variance plot to see number of components
```{r}
#cumulative scree plot
plot(cumsum(prop_varex), xlab = "Principal Component",
              ylab = "Cumulative Proportion of Variance Explained",
              type = "b")
```
We see that 30 components resulta a variance close to 90%

Use the 30 components 
```{r}
data2 <- data.frame(Alcoholic = mydata$STATE, prin_comp$x)

#1st 30 PCAs
data2 <- data2[,1:31]

#decision tree
library(rpart)
rpart.model <- rpart(Alcoholic ~ .,data = data2, method = "anova")
rpart.model

#transform tests into PCA
test.data <- predict(prin_comp, newdata = new_my_data)
test.data <- as.data.frame(test.data)
#30 components
test.data <- test.data[,1:30]

rpart.prediction <- predict(rpart.model, test.data)




```

Logistic regresion:
Data modification, calculate the mean of every row. It will be the explicative variable and STATE will be the answer (factor). That's X and Y.
```{r}
mydata.lr <-  mydata[2]
mydata.lr[2] <- c("0")
names (mydata.lr)[2] <- c("MEAN")

mydatanumeric <- mydata
mydatanumeric$USER <- NULL
mydatanumeric$STATE <- NULL
mydatanumeric$EXP_TYPE<- NULL

mymeans <- apply(mydatanumeric, 1, mean)

## filling the dataset with the MEAN values
for (i in 1:1729){
  mydata.lr[i,2] <- mymeans[i]
}

plot(mydata.lr$MEAN, jitter(mydata.lr$STATE, 0.15), pch = 19, xlab = "Means", ylab ="Alcoholic (0 - non, 1 - alcoholic)")
``` 
Alcoholic and non alcoholic trend to have the same means. So we can not define a model because there's no significative diference between alcoholic and non alcoholic.


