---
title: "Big Data Science - Assignment 2"
author: "Arnau Casanova & Alex Flores"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
subtitle: EEG Analysis
---

Load the dataset

```{r}
library(data.table)
library(magrittr)

mydata<-read.table("C:/Users/Arnau/Documents/BigDataScience/assignment1/bigTable.txt", header=TRUE )

```

Check availible variables (predictors)
```{r}
colnames(mydata)
```



Take a look at the structure

```{r}
mydata$STATE <-as.factor(mydata$STATE)
str(mydata)
```
Calculate the average of data and calculate the mean of the values
```{r}

#Convert categorical variables
library(dummies)
new_my_data <- dummy.data.frame(mydata, names = c("USER", "STATE", "EXP_TYPE"))
str(new_my_data)
```
PCA
```{r}
prin_comp <- prcomp(new_my_data, scale. = T)
names(prin_comp)
```
Mean and sd of variables
```{r}
prin_comp$center
prin_comp$scale
prin_comp$rotation
```
Principal component score vectors in 1729x91 dimension
```{r}
dim(prin_comp$x)
```
Plot the resultant principal compoents
```{r}
biplot(prin_comp, scale = 0)
```
Standard deviation of each principal compoent
```{r}
std_dev <- prin_comp$sdev
```
Variance
```{r}
pr_var <- std_dev^2
```
Check variance of first 10 components, higher value higher will be the information of these components
```{r}
pr_var[1:10]
```
Compute the proportion of variance explained by each component. Divide variance by sum total variance:
```{r}
prop_varex <- pr_var/sum(pr_var)
prop_varex[1:20]

```
1st component 31.9%, 2nd 15.33% and so on. How many components for the modeling? use scree plot to se variability in the data.
```{r}
#scree plot
plot(prop_varex, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             type = "b")
```
So 30 components explains around 98%variance in the data set. We had 67 predictors at the start, now 30.
Check the cumulative variance plot to see number of components
```{r}
#cumulative scree plot
plot(cumsum(prop_varex), xlab = "Principal Component",
              ylab = "Cumulative Proportion of Variance Explained",
              type = "b")
```
We see that 30 components resulta a variance close to 90%

Use the 30 components 
```{r}
data2 <- data.frame(Alcoholic = mydata$STATE, prin_comp$x)

#1st 30 PCAs
data2 <- data2[,1:31]

#decision tree
library(rpart)
rpart.model <- rpart(Alcoholic ~ .,data = data2, method = "anova")
rpart.model

#transform tests into PCA
test.data <- predict(prin_comp, newdata = new_my_data)
test.data <- as.data.frame(test.data)
#30 components
test.data <- test.data[,1:30]

rpart.prediction <- predict(rpart.model, test.data)


```
